\let\textcircled=\pgftextcircled
\chapter{Conclusions}
\label{chap:conclusions}

\par The project could be continued and a lot more Natural Language techniques could be applied to the models. The main constraint of this project has been time --- a lot of fascinating topics have only been touched slightly. A lot more time and other resources could be spent on this project. For example, noun phrase extraction has been tried as a part of the sentiment analysis, and it would be very useful to have it implemented. However, the scope of this project only allowed a library (\texttt{NLTK}) method to be used for noun phrase extraction, and it proved not to be very successful - each noun extraction would take almost 3 seconds, and a lot of noun phrases would not be detected. It would be beneficial to implement a fast and light noun phrase extractor, however, it would require a lot more time. 

The project leaves a lot of open questions that have not been answered by most recent research. For example, the Bag-of-Words approach is currently one of the most popular structures used in NLP. This is the simplest approach that allows for knowing almost nothing about the data that is being used. However, it would be interesting to see how the models that use the existing features of the data would perform. For example, is there a correlation between the location of the user and the sentiment of their posts? How can a tool detect idioms efficiently? 

Another unanswered question of NLP is whether irony and sarcasm could ever be detected by a machine, especially when people are not that good at detecting them themselves. 

Also, in the future work, comparing models created from the entire words to models created from only stems of the words and also maybe models created purely from affixes of the data could give unexpected results and be useful not only in Natural Language Processing but also in linguistics and literature. This research would show how much of emotional information different morphemes hold.

On a higher level, it might be valuable to understand how much emotional information different parts of speech hold. Intuitively, most of the information about the sentiment is being passed by adjectives and adverbs. Is it then possible to predict the sentiment of a sentence correctly using only words of these two classes? Would a model that contains no nouns perform much worse, and, if yes, how much worse?

If this project is ever continued, the person working on it should take all the points above into consideration. Despite the fact that the direction of the project had to be changed slightly, I hope that, with all the effort that has been put into this project, it will prove useful and contribute to a very exciting project and, hopefully, still move the Falkirk Council's project closer to the goal. 

It was incredibly engaging to work on this project. I learnt an immense amount of information, realised that NLP is one of the most interesting fields of Computer Science (for me, personally), and received a lot of experience in Natural Language Processing, Machine Learning, and working with a wide variety of APIs, had an amazing time working on the project. I also consider everything I have learnt throughout working on this project to be my largest achievement to date. 